
---
title: "Metabolomics Biomarker Analysis Report"
author: "Generated by R"
date: "`r Sys.Date()`"
output: 
  pdf_document:
    toc: true
    toc_depth: 2
    number_sections: true
geometry: margin=1in
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, 
                      fig.width = 8, fig.height = 6, dpi = 300)
```

# Executive Summary

This report presents the results of machine learning analysis for metabolomics biomarker discovery.

## Key Findings

```{r summary-table}
summary_data <- data.frame(
  Metric = c("Best Model", "Test Accuracy", "Test Kappa", "Number of Features", "Best Threshold"),
  Value = c("Ensemble", 
           round(results$results$individual_results$Ensemble$accuracy, 3),
           round(results$results$individual_results$Ensemble$kappa, 3),
           length(results$selected_features),
           results$results$best_threshold)
)
knitr::kable(summary_data, caption = "Analysis Summary")
```

# Model Performance

## Model Comparison

```{r model-comparison, fig.cap="Performance comparison across different machine learning models"}
create_model_comparison_plot(results)
```

```{r performance-table}
perf_table <- create_model_performance_table(results)
knitr::kable(perf_table, caption = "Detailed Model Performance Metrics")
```

# Biomarker Discovery

## Top Metabolite Biomarkers

```{r importance-table}
importance_table <- create_feature_importance_table(results, top_n = 10)
knitr::kable(importance_table[,c("Rank", "Metabolite_Clean", "Relative_Importance")], 
             caption = "Top 10 Metabolite Biomarkers by Importance", 
             col.names = c("Rank", "Metabolite", "Relative Importance (%)"))
```

## Partial Dependence Analysis

The following plots show how each metabolite concentration affects the predicted probability of the outcome.

```{r partial-dependence, fig.height=10, fig.cap="Partial dependence plots for top 4 metabolites"}
pd_plots <- create_partial_dependence_plots(results, top_features = 4)
grid.arrange(pd_plots[[1]], pd_plots[[2]], pd_plots[[3]], pd_plots[[4]], ncol = 2)
```

# Model Evaluation

## Prediction Distribution

```{r prediction-dist, fig.cap="Distribution of prediction probabilities by true class"}
create_prediction_distribution_plot(results)
```

## Confusion Matrix

```{r confusion-matrix}
cm_tables <- create_confusion_matrix_table(results)
knitr::kable(cm_tables$metrics, caption = "Classification Performance Metrics")
```

# Feature Relationships

## Metabolite Correlations

```{r correlation-heatmap, fig.height=8, fig.cap="Correlation matrix of selected metabolites"}
create_feature_correlation_plot(results)
```

# Conclusions

The ensemble model achieved **`r round(results$results$individual_results$Ensemble$accuracy*100, 1)`%** accuracy with a Kappa score of **`r round(results$results$individual_results$Ensemble$kappa, 3)`**, indicating `r ifelse(results$results$individual_results$Ensemble$kappa > 0.6, "substantial", ifelse(results$results$individual_results$Ensemble$kappa > 0.4, "moderate", "fair"))` agreement beyond chance.

The top biomarker metabolites identified were:
`r paste("- ", head(importance_table$Metabolite_Clean, 5), collapse = "\n")`

These results suggest promising metabolite biomarkers for further validation studies.

